\begin{itemize}
\item Vehicles and Points
\begin{itemize}
\item[CP] is the critical point
\item[SV] is the subject vehicle. The vehicle making a lane change.
\item[CL] is the vehicle in front of the SV in the SV's current lane.
\item[TL] is the vehicle in front of the SV in the target lane.
\item[LV] is the lag vehicle behind SV in the target lane.
\end{itemize}
\item Positions
\begin{itemize}
\item[$x$] is the unknown position of SV
\item[$x_l$] is the unknown position of LV
\item[$x_{cl}$] is the trajectory of CL
\item[$x_{tl}$] is the trajectory of TL
\item[$x(t)$] is the position of SV at time $t$
\item[$x_{cl}(t)$] is the position of CL at time $t$
\item[$x_{tl}(t)$] is the position of TL at time $t$
\item[$x_l(t)$] is the position of LV at time $t$
\item[$x_{clb}$] is the CL bound. SV must be located behind this bound during the lane change.
\item[$x_{tlb}$] is the TL bound. SV must be located behind this bound during the lane change.
\item[$x_{lb}$] is the LV bound. SV must be located in front of this bound during the lane change.
\item[$x_{min}(t)$] is the minimum of $x_{tlb}$ and $x_{clb}$ at time $t$
\end{itemize}
\item Times
\begin{itemize}
\item[$t_{end}$] is the available time until the 'string leader' reaches the critical point. The string is the set of cars involved in changing lanes.
\item[$0,t_{end}$] is the time interval available for the lane change manoeuvre.
\item[$\hat{t}$] is the earliest possible time a lane change of SV can be performed.
\item[$\Delta_{LC}$] is the time taken to change lanes.
\item[$W$] is the set of time windows where the leader vehicles enabled a lane change. It is given as 
\begin{equation}
W = {[t_1,t_2]|t_2 \ge t_1 + \Delta_{LC} \land v_{cl}(t) = v_{tl}(t) = v_{nom} \forall t \in [t_1,t_2]}
\end{equation}
\end{itemize}
\item Velocities
\begin{itemize}
\item[$v_{nom}$] is the nominal speed of the vehicles.
\item[$v_{up}$] is faster than average velocity.
\item[$v_{dn}$] is slow than average velocity.
\item[$v_{cl}$] is the CL velocity.
\item[$v_{tl}$] is the TL velocity.
\item[$v_{min}(t)$] is the minimum of $v_{tlb}$ and $v_{clb}$ at time $t$
\end{itemize}
\end{itemize}
